General Game Playing and Monte-Carlo Tree Search

### 2013-11-15 doc/paper/paper.tex
formatting mostly

\subsubsection{General Game Playing}
The General Game Playing Competition is an annual competition created and managed by Stanford University in which game-playing agents compete playing games of which they have no prior knowledge.\cite{StanfordGGP} This means they cannot use any heuristical knowledge (unless they can infer some from the game description) and must use more general methods of game-playing. Games occur with two or more agents communicating synchronously with a Game Master. The Game Master first tells the agents the rules of the game, and then for the rest of the game it is responsible for receiving player input and telling the players the next state of the game. All moves are synchronous, but turn-based games are easily simulated by the game rules requiring all but the active player to pass.

All the games played at the competition are described in Game Description Language (also known as Game Definition Language.) Games are boiled down to their most general essence when expressed in this language. A game description is composed of a series of relations. While the game can define its own game-specific relations, there are set relations that must be defined for all games. They are listed below, as described on the Stanford General Game Playing website \cite{StanfordGGP}.
\begin{itemize}
%TODO: italicize relations
\item role(a) means that a is a role in the game.
\item base(p) means that p is a base proposition in the game.
\item input(r,a) means that a is an action for role r.
\item init(p) means that the proposition p is true in the initial state.
\item true(p) means that the proposition p is true in the current state.
\item does(r,a) means that player r performs action a in the current state.
\item next(p) means that the proposition p is true in the next state.
\item legal(r,a) means it is legal for role r to play action a in the current state.
\item goal(r,n) means that player the current state has utility n for player r.
\item terminal means that the current state is a terminal state.
\end{itemize}

The true and does relations are in effect inputs. The true relations are inputs to the agents from the Game Master, and the does relation is the input from the agents to the Game Master. Therefore, all of the above relations except those two should be defined (most likely in terms of true and does) in the GDL description of a game.

@@

@@

@@

\subsubsection{Clojure}
The implementation of General Game Playing Monte-Carlo Tree Search presented in this paper was implemented in the programming language Clojure. Clojure is a Lisp dialect that can be compiled to run on the Java Virtual Machine, .NET Common Language Runtime, and JavaScript Virtual Machine. It was chosen for a variety of reasons
\begin{itemize}
% TODO: BOLD TITLE TERMS
\item Data Structures Clojure has a rich collection of immutable, persistent data structures including lists, vectors, sets, and maps.
\item Libraries Clojure has multiple libraries suited for the project, including pattern matching and logic programming libraries.
\item Highly Functional Clojure has a large set of higher-order functions that can operate on any of its collections.
\item Homoiconicity Part of this project involves translating Game Description Language into Clojure. This task would not have been possible in the timespan it was done in in a language that is not code-as-data.
\item Comfort On a personal level, I am most comfortable programming in Clojure.
\end{itemize}
\subsubsection{core.logic}
%FIXME: CAN'T LINEBREAK AFTER THIS SUBSUBSECTION
The primary Clojure library used in the implementation was core.logic, which is a port of the logic programming language miniKanren as formulated by William Byrd in his 2010 PhD thesis \cite{byrd2010relational}. miniKanren and core.logic are logic programing languages, which means that they deal in relations rather than functions. To get a feel for what that means, consider the + Clojure function compared to the + core.logic relation:

@@

(Note: The (run* [q] \textless RELATION\textgreater) means 'give me all the values of q such that the relation (with q substituted in) succeeds')

There are some key differences:
\begin{itemize}
\item The + function takes two arguments, a and b, and returns a+b, whereas the + relation takes three arguments, a b and c, and is the declaration that a+b=c.
\item This difference allows the relation to run backwards (in effect, performing subtraction.)
\item The relation doesn't run a number. It returns a list of numbers.
\end{itemize}

core.logic also allows the user to declare 'fresh' variables (which are free until bound) to work with. It also provides a relational disjunction in the form of conde. Finally, it offers unification ('==') to constrain logic variables. To show these in action (and to show a more complex example of core.logic), consider the below relation and its evaluation:

@@

While the core.logic version is more verbose, each part has a direct correspondence(SP?) to part of the GDL description:
\begin{itemize}
\item The relation has an explicit declaration with argument names.
\item All the reflexive relations at the head of each GDL \textless= block are transformed into a series of unifications with the corresponding GDL arguments. 
\item All free variables in the GDL relation are now explicitly declared in a core.logic fresh block.
\item The three separate GDL relations are now combined in a single conde block. In the GDL description, there is an implicit disjunction between the relations. conde is a core.logic disjunction.
\end{itemize}

>>>>

\subsubsection{General Game Playing}
The General Game Playing Competition is an annual competition created and managed by Stanford University in which game-playing agents compete by playing games of which they have no prior knowledge.\cite{StanfordGGP} This means they cannot use any heuristical knowledge (unless they can infer some from the game description) and must use more general methods of game-playing. Games occur with two or more agents communicating synchronously with a Game Master. The Game Master first tells the agents the rules of the game, and then for the rest of the game it is responsible for receiving player input and telling the players the next state of the game. All moves are synchronous, but turn-based games are easily simulated by the game rules requiring all but the active player to pass.

All the games played at the competition are described in Game Description Language (also known as Game Definition Language.) Games are boiled down to their most general essence when expressed in this language. A game description is composed of a series of relations. While the game can define its own game-specific relations, there are set relations that must be defined for all games. They are listed below, as described on the Stanford General Game Playing website \cite{StanfordGGP}.
\begin{itemize}
\item \emph{role(a)} means that a is a role in the game.
\item \emph{base(p)} means that p is a base proposition in the game.
\item \emph{input(r,a)} means that a is an action for role r.
\item \emph{init(p)} means that the proposition p is true in the initial state.
\item \emph{true(p)} means that the proposition p is true in the current state.
\item \emph{does(r,a)} means that player r performs action a in the current state.
\item \emph{next(p)} means that the proposition p is true in the next state.
\item \emph{legal(r,a)} means it is legal for role r to play action a in the current state.
\item \emph{goal(r,n)} means that player the current state has utility n for player r.
\item \emph{terminal} means that the current state is a terminal state.
\end{itemize}

The true and does relations are in effect inputs. The true relations are inputs to the agents from the Game Master, and the does relation is the input from the agents to the Game Master. Therefore, all of the above relations except those two should be defined (most likely partially in terms of true and/or does) in the GDL description of a game.

@@

\subsubsection{Summary}

\subsubsection{Critique}

@@

\subsubsection{Summary}

\subsubsection{Critique}

\subsection{Synthesis}
% Maybe I won't be able to write more than a few sentences
% In which case, remove the Synthesis section

@@

\subsubsection{Clojure}
The implementation of General Game Playing Monte-Carlo Tree Search presented in this paper was implemented in the programming language Clojure. Clojure is a Lisp dialect that can be compiled to run on the Java Virtual Machine, .NET Common Language Runtime, and JavaScript Virtual Machine. It was chosen for a variety of reasons
\begin{itemize}
% TODO: BOLD TITLE TERMS
\item \emph{Data Structures} Clojure has a rich collection of immutable, persistent data structures including lists, vectors, sets, and maps.
\item \emph{Libraries} Clojure has multiple libraries suited for the project, including pattern matching and logic programming libraries.
\item \emph{Highly Functional} Clojure has a large set of higher-order functions that can operate on any of its collections.
\item \emph{Homoiconicity} Part of this project involves translating Game Description Language into Clojure. This task would not have been possible in the timespan it was done in in a language that is not code-as-data.
\item \emph{Comfort} On a personal level, I am most comfortable programming in Clojure.
\end{itemize}
\subsubsection{core.logic}
%FIXME: CAN'T LINEBREAK AFTER THIS SUBSUBSECTION
The primary Clojure library used in the implementation was core.logic, which is a port of the logic programming language miniKanren as formulated by William Byrd in his 2010 PhD thesis \cite{byrd2010relational} and by Dan Friedman in his book The Reasoned Schemer \cite{reasonedschemer}.

miniKanren and core.logic are logic programing languages, which means that they deal in relations rather than functions. To get a feel for what that means, consider the + Clojure function compared to the + core.logic relation:

@@

(Note: The (\texttt{run* [q] \textless RELATION\textgreater)} macro means 'give me all the values of q such that the relation (with q substituted in accordingly) succeeds')

There are some key differences:
\begin{itemize}
\item The + function takes two arguments, a and b, and returns a+b, whereas the + relation takes three arguments, a b and c, and is the declaration that a+b=c.
\item This difference allows the relation to run backwards (in effect, performing subtraction.)
\item The relation doesn't return a number. It returns a list of numbers.
\end{itemize}

core.logic also allows the user to declare \texttt{fresh} variables (which are free until bound) to work with. It also provides a relational disjunction in the form of \texttt{conde}. Finally, it offers unification (\texttt{==}) to constrain logic variables. To show these in action (and to show a more complex example of core.logic), consider the below relation and its evaluation:

@@

While the core.logic version is more verbose, each part has a direct correspondence(SP?) to part of the GDL description:
\begin{itemize}
\item The relation has an explicit declaration with argument names.
\item All the reflexive relations at the head of each GDL \textless= block are transformed into a series of unifications with the corresponding GDL arguments. 
\item All free variables in the GDL relation are now explicitly declared in a core.logic \texttt{fresh} block.
\item The three separate GDL relations are now combined in a single \texttt{conde} block. In the GDL description, there is an implicit disjunction between the relations, and \texttt{conde} is a core.logic disjunction.
\end{itemize}
